\propertitle

### 2019-11-25 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-25T17:35:14+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
\caption{***}
\label{fig:plots_cellsubsets}
\end{figure}
% maxent_solutions_different_moments_partial.nb
These results show the potential usefulness of this maximum-entropy
approach. Here is a \emph{toy} example of possible uses. We could be
interested in the hypothesis that two distinct cell types or assemblies be
present in the region where the recording was made. Finding a
full-population distribution such as that in the upper panel of
\fig~\ref{fig:plots_5mom} would provide some evidence for this hypothesis.
Let's further imagine that we have reasons for suspecting that a specific
set of the sampled neurons are of one type, and the remaining of the other
type. In the case of data set 1, 27 of the 65 sampled neurons were
identified as grid cells belonging to 3--4 functional modules, see
\cite{dunnetal2015}. Could the two peaks in the distribution of
\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
cells? We apply the method to these two sets of neurons individually, using
$N=4\,150$ for the grid set and $N=5\,850$ for the non-grid set, to reflect
their proportions in the recorded sample. The results are shown in the
upper panel of \fig~\ref{fig:plots_cellsubsets}. The distribution for the
larger population of grid cells (\textcolor{mygreen}{green triangles})
seems to have one, broad peak, close to the first peak of the
full-population distribution. The distribution for the larger population of
non-grid cells (\textcolor{myred}{red circles}) has two peaks instead,
roughly at the same positions as the full-population distribution
(\textcolor{myblue}{blue curve}) but closer in height. So it would seem
that the population of grid cells is contributing to the first peak of the
full population, but it isn't the sole contributor. We can also assess how
much the distributions for the two sets are independent. If they were
completely independent, the full-population distribution would be given by
their convolution:
\begin{equation}
  \label{eq:convolution}
  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=0}^{\yA}} P_{\text{grid}}(\yA')\;
  P_{\text{non-grid}}(\yA-\yA'). %, \qquad \yA \in \set{1,\dotsc,10\,000}
\end{equation}
However, the lower panel of
figure~\ref{fig:plots_cellsubsets} shows that such convolution
(\textcolor{black}{black diamonds}) is different from the full-population
distribution (\textcolor{myblue}{blue curve}): the two peaks of the former
are closer in position and height. This means that the population
distributions of grid and non-grid cells are \emph{not} independent:
knowledge of the activity of either set gives us some information about the
activity of the other.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-25T17:52:38+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
\caption{***}
\label{fig:plots_cellsubsets}
\end{figure}
% maxent_solutions_different_moments_partial.nb
These results suggest that this maximum-entropy approach could be very
useful. Here is a toy example of possible uses. We could be interested in
the hypothesis that two distinct cell types or assemblies be present in the
region where the recording was made. Finding a full-population distribution
with two peaks, as in the upper panel of \fig~\ref{fig:plots_5mom}, would
provide some evidence for this hypothesis. Let's further imagine that we
have reasons for suspecting that a specific set of the sampled neurons are
of one type, and the remaining of the other type. In the case of the first
data set, 27 of the 65 sampled neurons were identified as grid cells
belonging to 3--4 functional modules, see \cite{dunnetal2015}. Could the
two peaks in the distribution of \fig~\ref{fig:plots_5mom} reflect the
activities of grid versus non-grid cells? We apply the method to these two
sets of neurons individually, using $N=4\,150$ for the grid set and
$N=5\,850$ for the non-grid set, to reflect their proportions in the
recorded sample. The results are shown in the upper panel of
\fig~\ref{fig:plots_cellsubsets}. The distribution for the larger
population of grid cells (\textcolor{mygreen}{green triangles}) seems to
have one, broad peak, close to the first peak of the full-population
distribution. The distribution for the larger population of non-grid cells
(\textcolor{myred}{red circles}) has two peaks instead, roughly at the same
positions as the full-population distribution (\textcolor{myblue}{blue
  curve}) but closer in height. So it would seem that the population of
grid cells is contributing to the first peak of the full population, but it
isn't the sole contributor. We can also assess how much the distributions
for the two sets are independent. If they were completely independent, the
full-population distribution would be given by their convolution:
\begin{equation}
  \label{eq:convolution}
  P_{\text{full}}(\yA) =  \smashoperator{\sum_{\yA'=0}^{\yA}} P_{\text{grid}}(\yA')\;
  P_{\text{non-grid}}(\yA-\yA'). %, \qquad \yA \in \set{1,\dotsc,10\,000}
\end{equation}
However, the lower panel of
figure~\ref{fig:plots_cellsubsets} shows that such convolution
(\textcolor{black}{black diamonds}) is different from the full-population
distribution (\textcolor{myblue}{blue curve}): the two peaks of the former
are closer in position and height. This means that the population
distributions of grid and non-grid cells are \emph{not} independent:
knowledge of the activity of either set gives us some information about the
activity of the other.
